A general framework for learning planning action models

### 2018-07-09 doc/aij18/learning_task.tex
small updates to learning section aij18 paper

This section details the \FAMA approach for learning action models. The notion of learning task is defined in section \ref{learning_task} and the components of the compilation scheme are described in sections \ref{propositional_encoding} and \ref{conditional_effects}. Subsequently, the compilation approach is fully detailed in section \ref{compilation} and the last section presents some theoretical properties of the compilation scheme.

@@

Our {\em learning task} consists in learning classical planning action models by observing one or more agents acting in a world, which is defined by a {\em classical planning frame} $\Phi=\tup{F,A}$. The learning task is formalized by the pair $\Lambda=\tup{\mathcal{M},\tau}$:

@@

Ultimately, we can always assume that $\mathcal{M}$ will contain the predicates $\Psi$ and the headers of the actions models, either explicitly provided in $\mathcal{M}$ or inferred from $\tau$. A {\em solution} to a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is a set of action models $\mathcal{M}'$ that is compliant with the input models $\mathcal{M}$ and the observed plan trace $\tau$.

@@

\begin{footnotesize}
	\begin{verbatim}
	;;; State observation
	(clear A) (on A B) (ontable B) (handempty)
	\end{verbatim}
\end{footnotesize}

@@

\subsection{Propositional encoding of an action model}
\label{propositional_encoding}

@@

In this section we formalize the propositional encoding of an \strips\ action model. This encoding is at the core of the \FAMA compilation approach for addressing the learning task defined in section \ref{learning_task}.

@@

We define $\Psi_v$ as the set of predicates $\Psi$ parameterized with the {\em variable names} of $\Omega_v$ as arguments. $\Psi_v$  defines the elements that can appear in the preconditions and effects of the action models. In the {\em blocksworld} domain, this set contains eleven elements, $\Psi_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}. In more detail, for a given action model $\xi$, we define $\Psi_{\xi}\subseteq \Psi_v$ as the subset of elements of $\Psi_v$ that appear in $\xi$. For instance, $\Psi_{\tt stack}=\Psi_v$ whereas $\Psi_{\tt pickup}$={\small\tt\{handempty, holding($v_1$),clear($v_1$),ontable($v_1$),on($v_1,v_1$)\}} excludes the elements from $\Psi_v$ that involve $v_2$ because {\small\tt pickup} actions have arity one. The size of the space of possible \strips\ models for a given $\xi$ is $2^{2|\Psi_{\xi}|}$ (recall that negative effects appear as preconditions and that they cannot be positive effects, and also that a positive effect cannot appear as a precondition). For the {\em blocksworld}, $2^{2|\Psi_{\tt stack}|}=4194304$ while for the {\tt pickup} operator this number is only 1024.

@@

Conditional effects allow us to compactly define actions whose effects depend on the current state. Supporting conditional effects is now a requirement of the IPC~\cite{vallati:IPC:AIM2015} and many classical planners cope with conditional effects without compiling them away.

@@

\subsection{Compilation}
\label{sec:compilation}

In section \ref{sec:motivation}, we exposed our reasons to solve the learning task via compilation to planning. Our compilation scheme  builds upon the approach presented in \cite{aineto2018learning}. \FAMA comes up with a more general and flexible scheme able to capture any type of input plan trace.

Our approach to solving a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is to compile $\Lambda$ into a classical planning task $P_{\Lambda}$ with conditional effects. The intuition behind the compilation is that a solution plan $\pi_\Lambda$ to $P_{\Lambda}$ induces the output set of action models $\mathcal{M}'$ that solves $\Lambda$. Specifically, a solution plan $\pi_\Lambda$ serves two purposes:

\begin{enumerate}
\item {\bf To program the set of action models $\mathcal{M}'$}. $\pi_\Lambda$ comprises a plan {\em prefix} whose actions determine the predicates $p\in \Psi_{\xi}$ that belong to $pre(\xi)$, $del(\xi)$ and $add(\xi)$ for each $\xi\in\mathcal{M}$.
\item {\bf To validate the set of action models $\mathcal{M}'$}. $\pi_\Lambda$ also comprises a plan {\em postfix} whose actions target the recognition of the available plan trace $\tau$ with the programmed action models $\mathcal{M}'$.
\end{enumerate}

>>>>

This section details the \FAMA approach for learning action models. The notion of the learning task is defined in section \ref{learning_task} and the components of the compilation scheme are described in sections \ref{propositional_encoding} and \ref{conditional_effects}. Subsequently, the compilation approach is fully detailed in section \ref{compilation} and the last section presents some theoretical properties of the compilation scheme.

@@

Our {\em learning task} consists in learning classical planning action models by observing one or more agents acting in a world definable by a {\em classical planning frame} $\Phi=\tup{F,A}$. The learning task is formalized by the pair $\Lambda=\tup{\mathcal{M},\tau}$:

@@

Ultimately, we can always assume that $\mathcal{M}$ will contain predicates in $\Psi$ as well as the headers of the actions models, either explicitly provided in $\mathcal{M}$ or inferred from $\tau$. A {\em solution} to a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is a set of action models $\mathcal{M}'$ that is compliant with the input models $\mathcal{M}$ and the observed plan trace $\tau$.

@@

\begin{footnotesize}
	\begin{verbatim}
	;;; State observation
	(clear A) (on A B) (ontable B) 
	\end{verbatim}
\end{footnotesize}

@@

\subsection{A propositional encoding for \strips\ action models}
\label{propositional_encoding}

@@

In this section we formalize a propositional encoding of an \strips\ action model. This encoding is at the core of the \FAMA compilation approach for addressing the learning task defined in section \ref{learning_task}.

@@

We define $\Psi_v$ as the set of predicates $\Psi$ parameterized with the {\em variable names} of $\Omega_v$ as arguments. The set $\Psi_v$  defines the elements that can appear in the preconditions and effects of the action models. In the {\em blocksworld} domain, this set contains eleven elements, $\Psi_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}. In more detail, for a given action model $\xi$, we define $\Psi_{\xi}\subseteq \Psi_v$ as the subset of elements of $\Psi_v$ that can appear in $\xi$. For instance, $\Psi_{\tt stack}=\Psi_v$ whereas $\Psi_{\tt pickup}$={\small\tt\{handempty, holding($v_1$),clear($v_1$),ontable($v_1$),on($v_1,v_1$)\}} excludes the elements from $\Psi_v$ that involve $v_2$ because {\small\tt pickup} actions have arity one. The size of the space of possible \strips\ models for a given $\xi$ is $2^{2|\Psi_{\xi}|}$ (recall that negative effects appear as preconditions and that they cannot be positive effects, and also that a positive effect cannot appear as a precondition). For the {\em blocksworld}, $2^{2|\Psi_{\tt stack}|}=4,194,304$ while for the {\tt pickup} operator this number is only 1024.

@@

Our approach to learning action models is to compile this learning task into a classical planning task with conditional effects.  Conditional effects allow us to compactly define actions whose effects depend on the current state. Supporting conditional effects is now a requirement of the IPC~\cite{vallati:IPC:AIM2015} and many classical planners cope with conditional effects without compiling them away.

@@

\subsection{Compilation}
\label{compilation}

In section \ref{sec:motivation}, we exposed our reasons to solve the learning task $\Lambda=\tup{\mathcal{M},\tau}$ via compiling it into a classical planning task $P_{\Lambda}$. Our compilation scheme  builds upon the approach presented in \cite{aineto2018learning} but \FAMA comes up with a more general and flexible scheme able to capture any type of input plan trace.

The intuition behind the \FAMA compilation is that a solution plan $\pi_\Lambda$ to $P_{\Lambda}$ induces the output set of action models $\mathcal{M}'$ that solves $\Lambda$. Specifically, a solution plan $\pi_\Lambda$ serves two purposes:

\begin{enumerate}
\item {\bf To program the set of action models $\mathcal{M}'$}. $\pi_\Lambda$ comprises a plan {\em prefix} whose actions determine the predicates $p\in \Psi_{\xi}$ that belong to $pre(\xi)$, $del(\xi)$ and $add(\xi)$ for each $\xi\in\mathcal{M}$.
\item {\bf To validate the set of action models $\mathcal{M}'$}. $\pi_\Lambda$ also comprises a plan {\em postfix} whose actions target the validation of the available plan trace $\tau$ with the programmed action models $\mathcal{M}'$.
\end{enumerate}
